---
title: "Aides Drug"
author: "Cesar Guillen"
date: "Summer 2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r eval=FALSE}
triUmp <- triUmp %>%
mutate(across(where(is.character), as.factor))
```

```{r}
females <- filter(triUmp, gender == "F")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

As the chart tells us, almost 60 percent of females who took the drug that was supposed to prevent getting aids didn't work for them. And we see that those who took the placebo seems to have gotten less cases of aids. Which proves that this drug might not be effective for females and their protection of aids while using this drug. But we must realize that this is still a small population of females who got aids, and not the whole population. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment,data=females)
rowPerc(table1)
colPerc(table1)
```

Of the females who took the Aids treatment, almost 60 percent of those have gotten Aids. Whereas those who took the placebo and have gotten Aids is only 40 percent. This table seems to prove that the drugs is not preventing aids, but is more likely to be inviting it in the body.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The tests prove that their are differences between the placebo and the drug that was trying to fight off Aids with their differences between their p-values. The p-value for this possibility is 7.686e-05, which means that there is a 7 out of 10000 chance that taking the drug would not be effective to fight off aids. This experiment is proving to be working in reverse, where taking the placebo shows to be more effective than the actual treatment. What else could be said is that there is only a 7 out of 10000 chance that the drug would be shown to be better than taking the placebo. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

With all the information we are given with the charts and the tests that we ran, it reveals that this drug is not going top be effective to fight off the Aids virus. As shown in the fisher tests there is only going to be a 7 out of 10000 chance that the  drug will be shown to be better than the placebo given to the other participants. That there is proof that this drug is not going to be effective and won't be given to the general public. What the graphs conclude to is that those in the blue are the females who have been given the drug are at a higher chance of getting Aids than those who have taken the placebo. There was a relationship between the drug and the disease, but not the desired one that would prove to be able to show that the drug is effective. If one were to take this drug is more likely to get the virus. 